---
title: "Compare MPG between automatic and manual transmission cars"
author:
date:
output: html_document
---
In this report I demonstrated the process to isolate the fittest model for the `mtcars` dataset. Through the diagnosis of this model, I picked up three outliers **`r rownames(mtcars)[c(17, 18, 20)]`** from the dataset. The comparison of transimission types `am` on the miles per gallon `mpg` was evaluated from the reservation of outliers to the removal of outliers. At the initiation of this process, I called the package `MASS`.  
```{r loadRequired, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(MASS)
```
### Exploratory Data Analyses
Firstly I explored the correlations for each pair of variables in `mtcars` dataset. As the below plot presented, five variables are categorical variables.  
```{r plot, echo=FALSE, results='asis', fig.align='center'}
pairs(mtcars, panel = panel.smooth, col = 3 + mtcars$am)
```

The original `mtcars` dataset was duplicated to the set `mtcars1` and the five categorical variables were transformed into `factor`. The model included all variables was stored in `all`:

```{r FacInterpretatioon, echo=TRUE, results='markup'}
mtcars1 <- mtcars
mtcars1$cyl <- as.factor(mtcars1$cyl)
mtcars1$vs <- as.factor(mtcars1$vs)
mtcars1$am <- as.factor(mtcars1$am)
mtcars1$gear <- as.factor(mtcars1$gear)
mtcars1$carb <- as.factor(mtcars1$carb)
all <- lm(mpg ~ . , data=mtcars1)
```
### Variable Selection
The fittest model was decided by the smallest AIC computed by the stepwise algorithm. The smallest AIC was acquried through `both` directions. Three predictors `wt`, `qsec`, and `am` were selected in the fittest model.  
```{r fit, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
fit <- stepAIC(all, direction="both")
```
```{r fitsum, echo=FALSE, results='markup', warning=FALSE, message=FALSE}
summary(fit)$coef
```
### Diagnostics of Residuals
Below plots illustrated that the residuals of most observations match the normalization assumption in addition to `r rownames(mtcars)[c(17, 18, 20)]`. The plot comparing resesuals and leverages indicate the three observations contributed the least influence to the model. Therefore, I indicated `r rownames(mtcars)[c(17, 18, 20)]` as the outliers of `mtcars` dataset.  
```{r diag, echo=FALSE, results='asis', fig.align='center' }
layout(matrix(c(1,2,3,4),2,2))
plot(fit)
```
To decrease the uncertainty of our conclusion, I compared the MPG betIen automatic and manual transmission in use of t test. Both T tests with and without outliers shoId the significant result that the mannual cars perfomred better on MPG than the automatic cars. I set sigificant level at `.05`.   
With the reservation of outliers, the average MPG for automatic car was **`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE, mtcars1))[6]) , digits=2)`** and the average MPG for mannual car was **`r round( as.numeric(unlist(t.test(mpg ~ am,var.equal = TRUE, mtcars1))[7]) , digits=2)`**. The statistical value was t(**`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1))[2]) , digits=2)`**) = **`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1))[1]) , digits=2)`**, p value **`r if(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1))[3] < .05){ c("< .05") }else{c("> .05")}`**.  
Without the outliers, the average MPG for automatic car was **`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1[c(1:16,19,21:32),]))[6]) , digits=2)`** and the average MPG for mannual car was **`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1[c(1:16,19,21:32),]))[7]) , digits=2)`**. The statistical value was t(**`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1[c(1:16,19,21:32),]))[2]) , digits=2)`**) = **`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1[c(1:16,19,21:32),]))[1]) , digits=2)`**, p value **`r if(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1[c(1:16,19,21:32),]))[3] < .05){ c("< .05") }else{ c("> .05")}`**.  

### Appendix
**Bold words** are produced by the codes embeded in line. Below column shows two examples I used in this report.  
```
Through the diagnosis of this model, I picked up three outliers ** r rownames(mtcars)[c(17, 18, 20)] ** from the dataset.   
...the average MPG for automatic car was ** r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1))[6]) , digits=2) **...
```