---
title: "Compare MPG between automatic and manual transmission cars"
author:
date:
output: html_document
---
### Overview
In this report we demonstrated the process to isolate the fittest model for the `mtcars` dataset. Through the diagnosis of this model, we picked up three outliers **`r rownames(mtcars)[c(17, 18, 20)]`** from the dataset. The comparison of transimission types `am` on the miles per gallon `mpg` was evaluated from the reservation of outliers to the removal of outliers. At the initiation of this process, we called the package `MASS`.  
```{r loadRequired, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(MASS)
```
### Exploratory Data Analyses
Firstly we explored the correlations for each pair of variables in `mtcars` dataset. As the below plot presented, five variables are categorical variables.  
```{r plot, echo=FALSE, results='asis', fig.align='center'}
pairs(mtcars, panel = panel.smooth, main = "mtcars data",     col = 3 + mtcars$am)
```

The original `mtcars` dataset was duplicated to the set `mtcars1` and transformed the five categorical variables into the `factor` class. The model included all variables was summarized and stored in `all`:

```{r FacInterpretatioon, echo=TRUE, results='markup'}
mtcars1 <- mtcars
mtcars1$cyl <- as.factor(mtcars1$cyl)
mtcars1$vs <- as.factor(mtcars1$vs)
mtcars1$am <- as.factor(mtcars1$am)
mtcars1$gear <- as.factor(mtcars1$gear)
mtcars1$carb <- as.factor(mtcars1$carb)
all <- lm(mpg ~ . , data=mtcars1)
```
###Variable Selection
The fittest model was decided by the smallest AIC in use of the stepwise algorithm. The smallest AIC was searched through `both` directions. The fittest model has three predictors `wt`, `qsec`, and `am`.
```{r fit, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
fit <- stepAIC(all, direction="both")
```
```{r fitsum, echo=FALSE, results='markup', warning=FALSE, message=FALSE}
summary(fit)$coef
```
###Diagnostics of Residuals
Below plots showed that the residuals of observations mostly match the normalization assumption in addition to `r rownames(mtcars)[c(17, 18, 20)]`. The plot comparing resesuals and leverages indicate the three observations contributed the least influence to the model. Therefore, we indicated `r rownames(mtcars)[c(17, 18, 20)]` as the outliers of `mtcars` dataset.
```{r diag, echo=FALSE, results='asis', fig.align='center' }
layout(matrix(c(1,2,3,4),2,2))
plot(fit)
```
###Conclusion
We compared the MPG between automatic and manual transmission in use of t test. Both T tests with and without outliers showed the same result that the mannual cars perfomred better on MPG than the automatic cars.  
With the reservation of outliers, the average MPG for automatic car was **`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE, mtcars1))[6]) , digits=2)`** and the average MPG for mannual car was **`r round( as.numeric(unlist(t.test(mpg ~ am,var.equal = TRUE, mtcars1))[7]) , digits=2)`**. The statistical value was t(**`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1))[2]) , digits=2)`**) = **`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1))[1]) , digits=2)`**, p value **`r if(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1[c(1:16,19,21:32),]))[3] < .05){ c("< .05") }else{c("> .05")}`**.  
Without the outliers, the average MPG for automatic car was **`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1[c(1:16,19,21:32),]))[6]) , digits=2)`** and the average MPG for mannual car was **`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1[c(1:16,19,21:32),]))[7]) , digits=2)`**. The statistical value was t(**`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1[c(1:16,19,21:32),]))[2]) , digits=2)`**) = **`r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1[c(1:16,19,21:32),]))[1]) , digits=2)`**, p value **`r if(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1[c(1:16,19,21:32),]))[3] < .05){ c("< .05") }else{ c("> .05")}`**. 


###Appendix
**Bold words** are codes embeded in line. Below column shows two examples we used in this report.
```
Through the diagnosis of this model, we picked up three outliers ** r rownames(mtcars)[c(17, 18, 20)] ** from the dataset.   
...the average MPG for automatic car was ** r round( as.numeric(unlist(t.test(mpg ~ am, var.equal = TRUE,mtcars1))[6]) , digits=2) **...
```